My title

### 2024-11-23 paper/paper.qmd
added to data section, halfway through model

# Introduction

@@

-   `CPI`: This variable represents the Consumer Price Index for All Urban Consumers, for all items in the U.S. City Average, which measures the average change over time in the prices paid by urban consumers for a market basket of consumer goods and services. Recorded monthly, this variable is measured in an index form, for which the base period is 1982-1984 where the value of the index sat at 100. Change in CPI from last year's CPI is one of the most common measures used to calculate inflation. This variable is seasonally adjusted and looks at \~ 88% of the population. **(CITE 2) HERE FROM REFERENCES)**

-   `Interest Rate`: This variable refers to the Federal Funds Effective Rate, which is the interest rate at which depository institutions trade federal funds with each other overnight. This interest rate is the central interest rate in the US, influencing other interest rates such as the prime rate (what banks charge customers with higher credit ratings) and longer term rates (mortgages, loans, savings). This variable is recorded monthly, represented as a percentage, and is not seasonally adjusted. **(CITE 3) HERE FROM REFERENCES)**

-   `Average Wage`: This variable denotes the Average Hourly Earnings of All Employees in the Total Private Sector. It includes premium pay for overtime and late-shift work. This was chosen over wage rates as this measures the actual return to a worker for a set period of time. It excludes benefits, irregular bonuses, retroactive pay, and payroll taxes paid by employers. This variable is measured in US Dollars per hour, recorded monthly, and is seasonally adjusted. **(CITE 4) FROM REFERENCES)**

-   `House Price`: This variable captures the Media Sales Price for New houses Sold int he United States. The media was chosen over the average prices to remove the effects of any outliers. This variable is recorded monthly, measured in US dollars, and is not seasonally adjusted. **(CITE 5) FROM REFERENCES)**

-   `House Supply`: This variables measures the Monthly Supply of New Houses in the United States. Measured in 'month's supply,' this refers to the ratio of new houses for sale to new houses sold. The month's supply indicates the size of the new for-sale inventory compared to the number of new houses currently being sold. This variable is recorded monthly and is seasonally adjusted. **(CITE 6) FROM REFERENCES)**

-   `House Price Growth`: This variable represents the year-over-year percentage growth in the Median Sales Price for New Houses Sold in the United States. It is calculated by taking the difference between the median house price at the end of the year and the beginning of the year, dividing this difference by the value at the beginning of the year, and multiplying by 100 to express the result as a percentage. This highlights the annual rate of change in house prices, helping to track trends in housing market inflation or deflation.

-   `Wage Rate Growth`: This variable reflects the year-over-year percentage growth in the Average Hourly Earnings of All Employees in the Total Private Sector. The calculation follows the same approach as house price growth: the difference between wages at the end of the year and the beginning of the year is divided by the wage value at the start of the year, then multiplied by 100 to express the result as a percentage. This variable captures how wages change annually, offering insights into trends in earning power.

@@

@@

@@

>>>>

# Introduction {#sec-introduction}

@@

-   **`CPI`:** This variable represents the Consumer Price Index for All Urban Consumers, for all items in the U.S. City Average, which measures the average change over time in the prices paid by urban consumers for a market basket of consumer goods and services. Recorded monthly, this variable is measured in an index form, for which the base period is 1982-1984 where the value of the index sat at 100. Change in CPI from last year's CPI is one of the most common measures used to calculate inflation. This variable is seasonally adjusted and looks at \~ 88% of the population. **(CITE 2) HERE FROM REFERENCES)**

-   **`Interest Rate`:** This variable refers to the Federal Funds Effective Rate, which is the interest rate at which depository institutions trade federal funds with each other overnight. This interest rate is the central interest rate in the US, influencing other interest rates such as the prime rate (what banks charge customers with higher credit ratings) and longer term rates (mortgages, loans, savings). This variable is recorded monthly, represented as a percentage, and is not seasonally adjusted. **(CITE 3) HERE FROM REFERENCES)**

-   **`Average Wage`:** This variable denotes the Average Hourly Earnings of All Employees in the Total Private Sector. It includes premium pay for overtime and late-shift work. This was chosen over wage rates as this measures the actual return to a worker for a set period of time. It excludes benefits, irregular bonuses, retroactive pay, and payroll taxes paid by employers. This variable is measured in US Dollars per hour, recorded monthly, and is seasonally adjusted. **(CITE 4) FROM REFERENCES)**

-   **`House Price`:** This variable captures the Media Sales Price for New houses Sold int he United States. The media was chosen over the average prices to remove the effects of any outliers. This variable is recorded monthly, measured in US dollars, and is not seasonally adjusted. **(CITE 5) FROM REFERENCES)**

-   **`House Supply`:** This variables measures the Monthly Supply of New Houses in the United States. Measured in 'month's supply,' this refers to the ratio of new houses for sale to new houses sold. The month's supply indicates the size of the new for-sale inventory compared to the number of new houses currently being sold. This variable is recorded monthly and is seasonally adjusted. **(CITE 6) FROM REFERENCES)**

-   **`House Price Growth`:** This variable represents the year-over-year percentage growth in the Median Sales Price for New Houses Sold in the United States. It is calculated by taking the difference between the median house price at the end of the year and the beginning of the year, dividing this difference by the value at the beginning of the year, and multiplying by 100 to express the result as a percentage. This highlights the annual rate of change in house prices, helping to track trends in housing market inflation or deflation.

-   **`Wage Rate Growth`:** This variable reflects the year-over-year percentage growth in the Average Hourly Earnings of All Employees in the Total Private Sector. The calculation follows the same approach as house price growth: the difference between wages at the end of the year and the beginning of the year is divided by the wage value at the start of the year, then multiplied by 100 to express the result as a percentage. This variable captures how wages change annually, offering insights into trends in earning power.

-   `Affordability Index`: This variable measures the ratio of the Median Sales Price for New Houses Sold in the United States to the Average Hourly Earnings of All Employees in the Total Private Sector. It represents how many hours of work are required to afford a median-priced home. A higher value indicates reduced affordability, as workers would need to work more hours to buy a home.

@@

### Affordability Index (CLEAN THIS UP)

library(tidyverse)
library(ggplot2)

ggplot(analysis_data, aes(x = Date, y = `Affordability Index`)) +
  geom_line(color = "blue") +
  labs(x = "Year", y = "Affordability Index") +
  theme_minimal()
```

@@

```{r}
# Load necessary library
library(tidyverse)

# Fit a simple multiple linear regression model
model <- lm(
  formula = `Affordability Index` ~ `Interest Rate (%)` + `CPI (Index)`, 
  data = analysis_data
)

# Summarize the model
summary(model)

@@

We aim to model the \*\*Affordability Index\*\*, which represents the number of hours of work required to afford a house, as a function of key economic variables such as interest rates and inflation (measured by CPI).

$$y_i = y_i = \beta_0 + \beta_1 \cdot \mathrm{CPI}_i + \beta_2 \cdot \mathrm{House    Supply}_i + \epsilon_i
$$

Where:

-   $y_i$ is the affordability index for observation iii, which measures the ratio of house prices to average wa ges, indicating housing affordability.

-   $\beta_0$ is the intercept, representing the expected value of the affordability index when all predictors are zero.

-   $\beta_1$ captures the effect of the CPI (Consumer Price Index) on the affordability index, reflecting inflation.

-   $\beta_2$ captures the effect of the Interest Rate on the affordability index.

-   $\epsilon_i$ represents the error term, assumed to follow a normal distribution with mean 0, accounting for unexplained variability in the affordability index.
